Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-05-30 paper/paper2/paper.tex
Paper changes and Ca I stuff

\documentclass[twocolumn]{aastex631}
\usepackage{graphicx}
\usepackage{physics}
\usepackage[super]{nth}
\usepackage{enumitem}

@@

\begin{abstract}
    The advent of machine learning methods is revolutionizing 
    numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference. Machine learning applied to spectroscopy may be expected to infer fundamental stellar parameters at greater speed, accuracy, and/or precision than conventional techniques. Here we introduce a new scalable ML technique to infer $T_{\mathrm{eff}}$,
    $\log(g)$, and [Fe/H] parameters, and the extensible method holds promise for other properties such as [$\alpha$/Fe], C/O, and
    $f_{sed}$. Our method contrasts with traditional data-driven
    approaches by introducing hybrid machine 
    learning (HML), a semi-empirical approach 
    to treating spectra as differentiable spectral lines described in 
    \texttt{blase} \citep{blase}.  
    \authorcomment1{Say how many spectra you analyze in the ``PHOENIX subset''.  State that an outcome of this work is the precomputed training.}
    We combine traditional linear interpolation 
    and Bayesian optimization to create a manifold mapping between 
    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) to fundamental paramerts.  of 
    128,723 unique spectral lines across a subset of the PHOENIX
    synthetic model grid \citep{PHOENIX} to create the PHOENIX generator, allowing
    us to reconstruct spectra at any continuously valued point 
    lying within the domain of the PHOENIX subset, and even reconstruct 
    batches of spectra in pseudo-parallel fashion, with the speedup 
    exceeding 20 times once more than 60 spectra are reconstructed. 
    We then combine the PHOENIX generator with Bayesian optimization to 
    efficiently infer fundamental stellar parameters. This study is 
    however a proof of concept, as we have not yet included various
    features that would be necessary for a full-fledged tool for 
    spectroscopic inference, however we show that semi-empirical 
    HML methods are a viable, albeit fledgling alternative to 
    more traditional approaches.
\end{abstract}

%token commit by gully!

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars
that produce them. Spectra encode fundamental properties such as temperature, 
surface gravity, and chemical composition via their seemingly countless 
absorption lines. Extrinsic properties such as the stellar radial velocity or 
the projected equatorial rotation further transform spectra. Our observations 
of the spectra degrade the otherwise pure astronomical signals, and so the 
resolution the observing instrument and other instrumental properties limit 
our ability to extract fundamental properties perfectly. Stellar spectra 
represent extremely complex, high-dimensional data that has gone through 
multiple transformations before reaching our detectors.
\authorcomment1{I like that you mention high-dimensional.  It recommend expanding on the meaning of high dimensional.  ``High'' in which sense?}

The challenge lies in extracting as much information as possible from
a stellar spectrum to get back the stellar fundamental properties. 
Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines appear
effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. This motivates the 
development of a more universal, instrument-agnostic open-source framework that can apply broadly to a range of spectral observations with relatively little tuning.

Multiple efforts have been made in this direction, treating spectra in 
different ways. The standard practice is to treat the wavelength and flux as 
simply two data arrays and bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions. Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, simplifying the data-model comparision stage nto a tractable regression, such as \texttt{starfish} \citep{starfish}. However, \texttt{blase} takes 
an intermediate approach \authorcomment1{cite}. 
It treats spectra not as a set of pixels or a set of  eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles. Each of these approaches has tradeoffs, but one key advantage places blase in a new category: the ability to fit the non-linear spectral line parameters with automatic differentiation.  This autodiff paradigm positions blase a promising semi-empirical tool.\authorcomment1{consider citing exojax and wobble, which also use autodiff for spectra}
\authorcomment1{Say a bit more on what blase paper 1 limitations-- only demonstrated on a few spectra with hand-chosen fiducial models, not scaled out to the entire grid.}

Ideally this process could be end-to-end differentiable, with all spectral inference code living inside a single autodiffable machine learning framework, like \texttt{PyTorch} or \texttt{JAX}. Here we separate the problem into two pieces.  First, the scaling out of the blase method to over 1000 precomputed synthetic spectral model clones, yielding a downloadable cache of pretrained machine learning models with over xx0,000 spectral lines.  This product alone serves as a useful asset to the community, since it can serve as a lookup for individual lines. Second, we explore the properties of the manifold mapping between fundamental stellar properties and the underlying spectral lines.  Finally, we show how reconstructions of the spectra through the manifold can be adapted for end-to-end spectral inference.
Extensions to this underlying framework have the potential to be the best in class with further research and development.

We chose the widely-adopted PHOENIX synthetic spectral model grid \authorcomment{cite?} as the basis for this study. The framework can be straightforwardly extended to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit our scope to a subset of the PHOENIX grid.

\authorcomment1{whoops, I repeated this above.  Key idea is to put some numbers here.}
Our study is divided into three main parts. First, we preprocess the PHOENIX
subset and detect spectral lines with \texttt{blase}. Second, we interpolate
the spectral line parameters, enabling us to evaluate the PHOENIX grid at 
any point within the bounds of the subset. Lastly, we use Bayesian optimization to power our inference algorithm and test its performance. An overview of this process is presented in Figure 1.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure1}
    \caption{Overview of the process used in this study.}
\end{figure*}

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the
PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
the grid, with the parameter ranges given in Table 1.
\authorcomment1{state the number of spectra in the subset if you have not already done so.  optionally state how many teff, logg, fe/h spacings.}

\begin{table}[h!]
    \centering
    \begin{tabular}{lll}
        \hline
        \bf{Parameter} & \bf{Symbol} & \bf{Range}\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms and to 
    ensure a rectilinear parameter space in order to work with \texttt{scipy}'s
    \texttt{RegularGridInterpolator} \citep{scipy}. The wavelength limits in particular 
    roughly line up with that of the Habitable Zone Planet Finder (HPF) 
    spectrograph \citep{HPF}. This subset is comprised of 1314 individual spectra.}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
three-step preprocessing pipeline, identical to \authorcomment1{blase paper 1}.
\begin{enumerate}
    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum 
    is known, the according blackbody spectrum was divided out.
    \item \textit{Percentile Normalization}: The spectra were normalized by dividing
    them by their 99th percentile.
    \item \textit{Continuum Normalization}: The spectra were further normalized by 
    dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}
Mathematically, we can express the preprocessing as follows:
\begin{gather}
    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
\end{gather}
where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the 
original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ 
is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the 
$n^\mathrm{th}$ percentile function. Arithmetic operations between arrays 
are assumed to be elementwise in all notation from here on out.

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral lines properties rather than a list of fluxes. This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation. 
Four parameters were optimized: the 
line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
learning rate of 0.05 over 100 epochs \citep{adam}. In addition, we limited two custom 
parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
that determines the extent of the Voigt profile to evaluate, saving
computational resources by not evaluating negligible line wings. Prominence sets a
lower limit for the amplitude of detected lines, which also saves resources 
by disregarding shallow lines. In short, larger values for wing cut and 
smaller values for prominence both increase the accuracy of using
\texttt{blase}'s spectral reconstruction at the expense of an
increased computational cost. In addition, it should be mentioned that
\texttt{blase} uses the pseudo-Voigt approximation, which saves on 
computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. It is a 
weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voight profile implementation uses the following:
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
\end{gather}
\end{subequations}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and 
Gaussian profiles, respectively. Notice that we use $\mu'$ instead of $\mu$ 
in the formula. This is because \texttt{blase} optionally allows the line center to shift 
slightly during optimization, and it is this shifted center which is used in 
computation. The individual Voigt profiles are still indexed by $\mu$ for 
cross-model line identification, explained in the next section. Once optimization 
was complete, the list of identified lines, was saved to a so-called ``state dictionary'' a common representation for pre-trained machine learning models that can be stored and reopened from this static cache.  These are stored as the \texttt{.pt} file format for each of the $N=xx,000$ PHOENIX subset 
grid point \citep{pytorch}. The total disk space these files took up is 465 MB.\authorcomment1{Restate the number of grid points for good measure, }.

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
As previously mentioned, \texttt{blase} tunes the line centers of
detected lines. This means that from one PHOENIX spectrum to the next, the
same line could have a slightly different line center. Since the goal of this
study is to interpolate the properties of each line, we needed to identify the
presence of a particular line across the PHOENIX subset, associating the same line with every occurrence. We decided to do
this by using the line centers $\mu$ of the detected lines pre-optimization. 
Now with each spectral line indexed by $\mu$, we had four parameters to interpolate:
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Note that since we are dealing with 
the parameters of Voigt profiles, we can see in Equation 2 that even if the 
interpolation method is linear, a final spectral reconstruction will vary 
non-linearly in the flux dimension.

There is also a second issue.  Spectral lines were often
only detected in some spectra from the PHOENIX subset. In Figure 2, we show
that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of 
    a slice of the PHOENIX subset at solar metallicity. We can see that
    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
    between grid points increases from 100 K to 200 K.}
\end{figure*}

This missing-lines phenomenon can arise for three different reasons: Stellar atmospherse genuinely do not produce that line at detectable strength at the given temperature and surface gravity. Second, PHOENIX missed it, possibly accidentally.  Or third, our line-finding and line-association algorithms missed it.  Whatever the cause, these missing lines have immediate practical consequences.  In theory, interpolation schemes may be inaccurate in regions where 
a line does not appear, or could break algorithms that demand a rectilinear grid, such as \texttt{scipy}'s \texttt{RegularGridInterpolator}.

To solve this, we artificially
populated missing grid points with log-amplitudes of -1000, which retained
interpolator stability by not being infinity, but also essentially 
nullified the line when evaluated. An example of the appearance of missing 
sections in heatmaps where a line does not appear is shown in Figure 3. 
In total, across the entire PHOENIX subset, \texttt{blase} detected 
128,723 individual spectral lines.

Every one of these lines can be visualized as a manifold: a mapping between one 4-dimensional heatmap of line properties to a 3-dimensional heatmap of fundamental properties. All of these 128,723 manifolds have to be interpolated.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure3}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset
    slice at solar metallicity. Notice the missing chunk in the top left
    of the figure; \texttt{blase} did not detect a spectral line here,
    but we have to artificially populate those points with lines that 
    have $\ln(a) = -1000$. The change in spacing on the x-axis is owing to
    the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 
    200 K after $T_\mathrm{eff} = 7000$ K.}
\end{figure*}

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was then
written to disk in the \texttt{.pkl} file storage format.  These interpolators generate multiple 
manifolds with the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuously evaluable function, sometimes called a spectral emulator or surrogate model. With the given size of the PHOENIX subset, the 
interpolator list takes up 13.2 GB of disk space. This evaluation 
is able to reconstruct an existing PHOENIX spectrum or alternatively 
interpolate a new spectrum within the domain of the PHOENIX subset,
we call this the PHOENIX generator. In Figure 4, we show the same 
spectral line as in Figure 3, but now supersampled using the PHOENIX
generator evaluated over the same slice.

The spectral reconstruction process is done by iterating over the 
PHOENIX generator, evaluating each interpolator at the given coordinates, 
then reshaping the data into the same format that PyTorch uses for 
state dictionaries. During the iteration, if the interpolated
log-amplitude of the line is less than -100, the line is excluded from the 
state dictionary. The rationale for -100 instead of -1000 originates from the fact that at wavelength regions near artificially populated lines, the manifold experiences artifacts due 
a sudden decline. We have to anticipate that decline and cut off the 
behavior in advance. This saves on computational time. 
\authorcomment1{I do not understand this.  Maybe consider removing?}

Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
which reconstructs the spectrum by constructing a forward model based on 
the input state dictionary. Any \texttt{nan} values are set to 1 (which we 
can do because the spectra are all normalized), and the spectrum is 
complete.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure4}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX 
    subset slice at solar metallicity, now supersampled with the PHOENIX 
    generator. Notice that the missing chunk in the top left still exists 
    and does not display any artifacts, as the artificially populated points 
    are removed after interpolation to retain the model's integrity. Also
    see that the x-axis spacing is now uniform, as the PHOENIX generator 
    was evaluated at constant step.}
\end{figure*}

>>>>

\documentclass[twocolumn]{aastex631}
\usepackage{graphicx}
\usepackage{physics}
\usepackage[super]{nth}
\usepackage{enumitem}
\usepackage[all]{hypcap}

@@

\begin{abstract}
    The advent of machine learning (ML) methods has and continues to revolutionize numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference.
    ML methods in spectroscopy are expected to perform with greater efficacy than conventional computational techniques.
    In this study we introduce a hybrid ML (HML) method that combines automatic differentiation (autodiff), interpolation, and Bayesian optimization to infer stellar parameters given stellar spectra.
    The stellar parameters we study are $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method has potential for extension to other properties such as [$\alpha$/Fe], C/O, and $f_{sed}$.
    We first use the semi-empirical approach to treating spectra as sets of tunable spectral lines introduced in \texttt{blase} \citep{blase}, contrasting with traditional data-driven approaches. 
    \texttt{blase} is used on 1,314 spectra from the PHOENIX synthetic spectral model grid \citep{PHOENIX}, identifying 128,723 unique spectral lines.
    For each of these lines, a manifold mapping stellar parameters to spectral line parameters is created using regular grid linear interpolation.
    These manifolds are aggregated to to create the PHOENIX generator, sporting parallelized reconstruction of spectra at continuously valued points within the PHOENIX subset.
    Gaussian Process minimization is then used to infer stellar parameters by minimizing the root-mean-square (RMS) loss between the input spectrum and PHOENIX generator spectra.
    From testing, the inference error in $T_{\mathrm{eff}}$ was 185 K, $\log(g)$ was 0.19, and [Fe/H] was 0.12 dex.
    Our products are threefold: first an archive of the state dictionaries of the \texttt{blase} clones of the PHOENIX subset, second the PHOENIX generator itself, and third, the reconstruction and inference algorithm.
    This study represents a rudimentary proof of concept showing that semi-empirical HML is a viable, albeit fledgling alternative to traditional approaches.
\end{abstract}

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars that produce them. 
Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their seemingly countless absorption lines. 
Extrinsic properties such as the stellar radial velocity or the projected equatorial rotation further transform spectra. 
Our observations of the spectra degrade the otherwise pure astronomical signals, and so the resolution the observing instrument and other instrumental properties limit our ability to extract fundamental properties perfectly. 
Stellar spectra represent extremely complex, data, influenced by multiple parameters, that have gone through multiple transformations before reaching our detectors.

The challenge lies in extracting as much information as possible from a stellar spectrum to get back its stellar parameters. 
Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters from spectra (ASPCAP in the case of APOGEE), and these pipelines appear effective \citep{APOGEE, ASPCAP}. 
However, they all work on the core assumption that the data to be analyzed is a list of pixels. 
Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. 
This motivates the development of more universal, instrument-agnostic open-source frameworks that can apply broadly to a range of spectral observations with relatively little tuning.

Multiple efforts have been made in this direction, treating spectra in different ways. 
The standard practice is to treat the wavelength and flux as simply two arrays and use bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions. 
Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, implementing the data-model comparison stage as a tractable regression, such as \texttt{starfish} \citep{starfish}. 
However, \texttt{blase} takes an intermediate approach \citep{blase}, treating spectra not as a set of pixels or a set of eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles.
Each of these approaches has tradeoffs, but one key distinction places \texttt{blase} in another category defined by the ability to fit the nonlinear spectral line parameters with autodiff.
We see autodiff used in tools such as \texttt{exojax} and \texttt{wobble} \citep{exojax, wobble}, but the recasting of spectra into sets of inherently nonlinear spectral lines positions \texttt{blase} as a unique and promising semi-empirical tool.
The original \texttt{blase} paper demonstrated the ability to tune spectral lines with autodiff, but only on a few hand-picked fiducial models, not scaled out to an entire model grid.

Ideally this process could be monolithic, with the end-to-end spectral inference code powered by a single autodiffable machine learning framework, like PyTorch or JAX \citep{pytorch, jax}. 
However, here we separate the problem into three pieces. 
First, the scaling out of the blase method to 1,314 precomputed synthetic spectral model clones, yielding a downloadable archive of pretrained machine learning models with 128,723 unique spectral lines. 
Second, the creation of manifolds mapping stellar parameters to spectral line parameters using regular grid linear interpolation. 
Finally, we show how reconstructions of the spectra using said manifolds can be used for spectral inference.
An overview of this process is shown in \autoref{fig:figure1}.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure1}
    \caption{Overview of the process used in this study.}
    \label{fig:figure1}
\end{figure*}
Advancements to this underlying basis have the potential to be top tier options with further research and development.

We chose the widely-adopted PHOENIX synthetic spectral model grid \citep{PHOENIX} as the basis for this study.
Our approach can be straightforwardly applied to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, but for now we limit our scope to a subset of the PHOENIX grid.

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the PHOENIX synthetic spectral model grid. 
Instead, we focused on a subset of the grid, with details given in \autoref{tab:table1}.

\begin{table}[h!]
    \centering
    \begin{tabular}{lll}
        \hline
        \bf{Parameter} & \bf{Symbol} & \bf{Range}\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. 
    These limits were imposed to reduce the computational cost of the algorithms and to ensure a rectilinear parameter space in order to work with \texttt{scipy}'s \texttt{RegularGridInterpolator} \citep{scipy}. 
    The wavelength limits in particular roughly line up with that of the Habitable Zone Planet Finder (HPF) spectrograph \citep{HPF}. 
    This subset is comprised of 1,314 individual spectra: 73 unique $T_{\mathrm{eff}}$ values, 9 unique $\log(g)$ values, and 2 unique [Fe/H] values.}
    \label{tab:table1}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
three-step preprocessing pipeline similar to that from \citealt{blase}.
\begin{enumerate}
    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum is known, the according blackbody spectrum was divided out.
    \item \textit{Percentile Normalization}: The spectra were normalized by dividing them by their 99th percentile.
    \item \textit{Continuum Normalization}: The spectra were further normalized by dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}
Mathematically, we can express the preprocessing as follows:
\begin{gather}
    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
\end{gather}
where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the $n^\mathrm{th}$ percentile function. 
Arithmetic operations between arrays are assumed to be elementwise in all notation from here on out.

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral line properties rather than a list of fluxes. 
This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation. 
Four parameters were optimized: the line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the Lorentzian width $\gamma$. 
The optimization used the Adam optimizer with a learning rate of 0.05 over 100 epochs \citep{adam}. 
In addition, we limited two custom parameters: wing cut to 6000 and prominence to 0.005. 
Wing cut is a parameter that determines the extent of the Voigt profile to evaluate, saving computational resources by not evaluating negligible line wings. 
Prominence sets a lower limit for the amplitude of detected lines, which also saves resources by disregarding shallow lines. 
In short, smaller values for wing cut and smaller values for prominence both decrease the computational cost of \texttt{blase}'s cloning process at the expense of decreasing its accuracy slightly. 
In addition, it should be mentioned that \texttt{blase} uses the pseudo-Voigt approximation, which saves on computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. 
It is a weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voight profile implementation uses the following:
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
    \label{eq:equation2}
\end{gather}
\end{subequations}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and Gaussian profiles, respectively. 
Notice that we use $\mu'$ instead of $\mu$ in the formula. 
This is because \texttt{blase} optionally allows the line center to shift slightly during optimization, and it is this shifted center which is used in computation. 
The individual Voigt profiles are still indexed by $\mu$ for cross-model line identification, explained in the next section. 
Once optimization was complete, the list of identified lines, was saved to a `state dictionary': a common representation for pre-trained machine learning models that can be stored to disk for reuse later. 
These are stored in the \texttt{.pt} file format for each of the 1,314 PHOENIX subset grid points. 
The total disk space these files take up is 465 MB (382 MB when downloaded as a zip archive).

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
As previously mentioned, \texttt{blase} tunes the line centers of detected lines. 
This means that from one PHOENIX spectrum to the next, the same line could have a slightly different line center. 
Since the goal of this study is to interpolate the properties of each line, we needed to identify the presence of a particular line across the PHOENIX subset, associating the same line with every occurrence. 
We decided to do this by using the line centers $\mu$ of the detected lines pre-optimization. 
Now with each spectral line indexed by $\mu$, we had four parameters to interpolate: $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. 
Note that since we are dealing with the parameters of Voigt profiles, we can see in \autoref{eq:equation2} that even if the interpolation method is linear, a final spectral reconstruction will vary nonlinearly in flux.

There is also a second issue. Spectral lines were often only detected in some spectra from the PHOENIX subset. 
In \autoref{fig:figure2}, we show that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of 
    a slice of the PHOENIX subset at solar metallicity. We can see that
    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
    between grid points increases from 100 K to 200 K.}
    \label{fig:figure2}
\end{figure*}

This missing-lines phenomenon can arise for three different reasons: Stellar atmospherse genuinely do not produce that line at detectable strength at the given temperature and surface gravity. 
Second, PHOENIX missed it, possibly accidentally. 
Or third, our line-finding and line-association algorithms missed it. 
Whatever the cause, these missing lines have immediate practical consequences. 
Rectilinear interpolation schemes break in regions where a line does not appear; you can't interpolate a quantity that simply doesn't exist.

To solve this, we artificially populated missing grid points with log-amplitudes of -1000, which retained interpolator stability while nullifying the evaluated line. 
An example of the appearance of missing sections in heatmaps where a line does not appear is shown in \autoref{fig:figure3}. 
In total, across the entire PHOENIX subset, \texttt{blase} detected 128,723 individual spectral lines.  
Every one of these lines can be visualized as a manifold mapping a 3D stellar parameter vector to a 4D spectral line parameter vector, and every one must be interpolated.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure3}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity. 
    Notice the missing chunk in the top left of the figure; \texttt{blase} did not detect a spectral line here, but we have to artificially populate those points with lines that have $\ln(a) = -1000$. 
    The change in spacing on the x-axis is owing to the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 200 K after $T_\mathrm{eff} = 7000$ K.}
    \label{fig:figure3}
\end{figure*}

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], and the output was a list of four parameters, $\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. 
For each line, one of these interpolator objects was created using linear interpolation, and these interpolators were aggregated into a single list, which was then written to disk in the \texttt{.pkl} file storage format.  
These interpolators generate multiple manifolds representing the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset into a continuously evaluable function, sometimes called a spectral emulator. 
With the given size of the PHOENIX subset, the interpolator list takes up 13.2 GB of disk space. 
This evaluation is able to reconstruct an existing PHOENIX spectrum or alternatively interpolate a new spectrum within the domain of the PHOENIX subset, so we call this the PHOENIX generator. 
In \autoref{fig:figure4}, we show the same spectral line as in \autoref{fig:figure3}, but now supersampled using the PHOENIX generator evaluated over the same slice.

The spectral reconstruction process is done by iterating over the PHOENIX generator, evaluating each interpolator at the given coordinates, then reshaping the data into the same format that PyTorch uses for state dictionaries. 
During the iteration, if the interpolated log-amplitude of the line is less than -100, the line is excluded from the state dictionary. 
We do this to avoid artifacts in the manifolds due to the artifical population of log-amplitudes of -1000 where grid points were missing.

Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator}, which reconstructs the spectrum by constructing a forward model based on the input state dictionary. 
Any \texttt{nan} values are set to 1 (which we can do because the spectra are all normalized), and the spectral reconstruction is complete.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure4}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset slice at solar metallicity, now supersampled with the PHOENIX generator. 
    Notice that the missing chunk in the top left still exists and does not display any artifacts, as the artificially populated points are removed after interpolation to retain the model's integrity. 
    Also see that the x-axis spacing is now uniform, as the PHOENIX generator was evaluated at constant step.}
    \label{fig:figure4}
\end{figure*}

\section{Bayesian Inference and Testing}
\subsection{Spectral Reconstruction Time}
A typical use case for the PHOENIX generator may be to batch reconstruct spectra from an array of input stellar parameters. 
Therefore, there is some motivation to reduce the computational cost of this procedure to tractable levels. 
We evaluate the computational time needed to use the PHOENIX generator in two distinct ways. 
First, for a single input, which would be relevant in serial applications. 
Second, for an array of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of an entire array of input coordinates to be evaluated at once. 
However using \texttt{blase} to reconstruct the spectrum from our interpolated state dictionary is always done serially, leading to what is actually more of a psuedo-parallel evaluation, but extremely performant nonetheless. 
Performance results are shown in \autoref{fig:figure5}, and we can see that the multi-reconstruction is much faster than a series of single reconstructions.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure5}
    \caption{Line chart showing the time taken to reconstruct varying numbers of spectra using the PHOENIX generator (lower is better). 
    We can see that the time taken per spectra for the serial implementation hovers around 15 seconds within run-to-run variance, while the pseudo-parallel implementation continually decreases in time taken per spectrum as the number of inputs increases. 
    The speedup factor (higher is better) increases as more spectra are generated, which is also a desirable outcome.}
    \label{fig:figure5}
\end{figure}
